Internet of Things: a Real-World Deployment Towards Smart Grids Integration

### 2013-07-09 doc/thesis/report.tex
tiny modifications done

\usepackage{etoolbox}
\makeatletter
\patchcmd{\thebibliography}{%
  \chapter*{\bibname}\@mkboth{\MakeUppercase\bibname}{\MakeUppercase\bibname}}{%
  \chapter{Index}}{}{}
\makeatother

@@

\chapter{Abstract}
\paragraph{}
Information Technology is growing exponentially and largely impacting our daily lives. The current evolution in computing is the \gls{iot}. The Internet is smoothly evolving from an Internet of People to and Internet of \textit{Things}. By 2020, it is expected to have 50 billion \textit{Things} connected to the Internet. 
\paragraph{}
However, such a migration induces a strong level of complexity when handling interoperability between heterogeneous \textit{Things}, e.g. Wireless Sensors, RFID, and mobile devices.
\paragraph{}
Another migration is taking place in the power grid that is evolving towards the \gls{sg}. In fact, the Information Technology will play a key role in implementing needed functions such as \gls{dr}\glsreset{dr} and integrating Distributed and Renewable Energy Sources into \gls{sg}. The \gls{iot} is expected to enable a smooth integration of Distributed Renewable Energies within the electric grid.
\paragraph{}
This thesis focuses on the development of an \gls{iot} test-bed, using wireless sensors, that will be used for smart grid applications. The test-bed built provides IPv6 gatewaying, remote control of wireless sensor nodes which is built on top of \gls{6lowpan}\glsreset{6lowpan}, and mobile access to the system.
\paragraph{}
This study results in the creation of a test-bed that provides an infrastructure for integrating \gls{sg} into \gls{iot} and build the overlying services such as \gls{dr}. Therefore, the system's performance was put to test by conducting two experiments where in each the delay and jitter were measured to make sure the system can handle overlying services in future projects.

\chapter*{Résumé}
\paragraph{}
Les nouvelles technologies de l'information connaissent une croissance exponentielle et prennent de plus en plus d'impact sur nos vies. L'évolution en cours de l'informatique est \glsreset{iot}\gls{iot}. L'Internet est discrètement en train de transitioner d'un réseau pour les personnes vers un reseau pour les \textit{Choses}. En 2020, il est prévu d'avoir 50 milliards \textit{Choses} connectées à Internet.
\paragraph{}
Cependant, une telle migration induit un niveau élevé de complexité lors de la manipulation de l'interopérabilité entre \textit{Choses} hétérogènes, par exemple, Capteurs sans fil, RFID, et appareils mobiles.
\paragraph{}
Une autre migration commence dans le réseau électrique qui est entrain de se transformer en un nouveau réseau appelé \glsreset{sg}\gls{sg}. Ce nouveau réseau compte principalement sur les avantages des technologies de l'information qui lui permettra d'intégrer de nouvelles fonctionnalités telle que \glsreset{dr}\gls{dr} et la l'exploitation décentralisée des énergies renouvelables. Pour cela, \gls{iot} va permettre une douce intégration des énergies renouvelables dans le réseau électrique.
\paragraph{} 
Cette thèse porte à développer un \gls{iot} prototype qui utilise des capteurs sans fil pouvant être utilisés dans le Smart Grid. Ce prototype fournit des services tels qu'une passerelle IPv6, le control à distance des capteurs sans fil, et un accès mobile au système.
\paragraph{}
Cette étude conduit à la création d'un prototype qui propose une infrastructure for intégrer \gls{sg} au sein de \gls{iot} et donc permettre de développer des services tel que \gls{dr}. Pour cela, les performances du système doivent être mises en test pour s'assurer de la possibilités de compter sur ce prototype. Dans ce contexte, deux expériences ont été menées mesurant le retard et le gigue pour savoir si le système peut supporter d'autres applications.
\setcode{utf8}

@@

\section{Importance of Study}
\paragraph{}
The global electrical grid is the basis that gives us this quality of life that most people around the world enjoy. The global electrical grid is sometimes called the largest machine ever made by man \cite{ref1}. It is a widespread and distributed machine that provides power to households, companies and organizations. Therefore the whole world's economy depends on it. The electrical grid knows numerous problems such as blackouts, and current instabilities that cost billions of dollars yearly to both the utilities and insurances. The grid suffers from the problem of peak hours where energy consumption is at its highest level and that one single utility cannot provide.
\paragraph{}
The \glsreset{sg}\gls{sg} comes as a solution to the problems of the traditional electrical grid. \gls{sg} is an intelligent, auto-balancing, self-monitoring, and self-healing electrical grid \cite{ref2} and whose energy sources are distributed between utilities and consumers. The nature of the \gls{sg} as having distributed energy sources, reduces numerous losses that arise from the long distance transmission lines, and also from the stress and the overload of the electrical lines.
\paragraph{}
Another revolution began, and it is taking place in the Information Technology. It is called the \glsreset{iot}\gls{iot}. \gls{iot} is the era of autonomous machine being able to produce and consume information, and being able to act autonomously without the guidance of humans, to perform their tasks. Cisco predicts that the number of devices to be connected to the Internet in 2020 will break the 50 billions devices from 500 million in 2003. That would make 6.58 devices connected for each person in the world whereas that same ratio was 0.07 17 years earlier \cite{ref3}. This clearly shows that the shift to the \gls{iot} will be exponential.
\paragraph{}
\glsreset{sg}\gls{sg} should benefit from the \gls{iot}. The integration of the \gls{sg} into the \gls{iot} constitutes the main topic of this study and goes along with increasing interest recently raised in both academia and industry, on the inherent correspondence between \gls{sg} and \gls{iot}. This  correspondence stems mainly from the need to disseminate information between the \gls{sg} components in order to make optimal decisions about energy use. This information dissemination could only be optimal when deploying ubiquitous sensoring and metering, and using the available network infrastructure, i.e. the Internet.

\section{Rationale of Study}
\paragraph{}
This project has been done in order to investigate ways the \gls{iot} can benefit the \gls{sg}. The Smart Grid needs a set of novel functionalities. One of these ideas is the two-way communication that will make it possible for the utility to communicate with consumers and vice versa. This is possible in the Internet Of Things as devices in \gls{iot} communicate by producing and consuming information. Numerous ideas assume the existence of the two-way communication such as the \gls{dr} which offers a set of services for both the power provider and the consumer.
\paragraph{}
The \glsreset{iot}\gls{iot} uses different technologies such as \gls{wsn}, \gls{rfid}, \gls{ipv6} and \gls{6lowpan}\glsreset{6lowpan} which makes the Internet of Things a very heterogeneous and complex system \cite{ref4}. One of the challenges in \gls{iot} is to deal with this heterogeneity in order to make the communication smooth and transparent to the parties involved. This heterogeneity should become invisible to any two devices willing to communicate in the \gls{iot}.

\section{Problem Statement}
\paragraph{}
The \glsreset{sg} \gls{sg} relies on the two-way communication to build a whole communication framework and a protocol stack namely the \glsreset{dr}\gls{dr} and other services built on top of this last. From the consumer's/household's perspective or more precisely the household's side,there are smart home energy management systems in the market nowadays, but none of them offers an integration with \gls{sg}. To have such an integration with \gls{sg}, there is a need for the two-way communication, in order to make it possible for any appliance to report its real-time energy consumption, and to be able to be controlled remotely. Such pillar is missing for the time being, and therefore the integration of the smart home within the smart grid is not possible. Also, in order for \gls{sg} to be integrated into the \gls{iot}, all devices should be able to communicate in the two ways.
\paragraph{}
\gls{sg} counts on various technologies to achieve its main goals. This use of different technologies comes with a problem which is the heterogeneous nature of the system. As numerous components count on different technologies, there is a need to interface them in order to make this heterogeneity transparent to the all components of the system. As a matter of fact, a software should be built within the system whose main goal is to take care of making the communication technology-transparent. This component will give the possibility to all components of the system to work hand-in-hand in order to achieve their goals independently of the data link layer (Ethernet, Wifi or Zigbee) or the network layer technology (\gls{ipv4} or \gls{ipv6}) they are using.

\section{Purpose of Study}
 \paragraph{}
The goal of the study is to solve the problems introduced in the problem statement section. This means creating the two-way communication, and solving the heterogeneity that arises from integrating smart homes with the smart grid. In fact, each components in the smart home will become part of the Internet of Things by making it communicate with any device in the Internet by having an \gls{ipv6} address using \gls{6lowpan} whose goal is to provide the TCP/IP to the smallest devices and also those with the least computational resources.
\paragraph{}
The first purpose of this study is to create a testbed having the two-way communication and also being able to make use of it in order to build richer applications based on it. To do so, \gls{6lowpan} would had to be integrated within every device in the \gls{wsn}. This would make every sensor node (mote) identifiable in the network, ready to communicate and be part of the \gls{iot}. This communication will go in both directions and based on \gls{6lowpan}; the two-way communication will be built.
\paragraph{}
The second purpose of this study is to solve the heterogeneity problem that arises from the integration of smart homes into the grid and also when making the whole system part of the \glsreset{iot}\gls{iot}. Solving such a problem means having an extra component that will track all  packets and change them to fit the standards of the targeted component. This will make of this additional component a middleman and that would make the heterogeneity problem disappear.

\section{Scope of Study}
\paragraph{}
This study aims at solving the two problems presented in section 1.3. To do so, a variety of tools and technologies were used for this purpose.The two problems could not be solved one independently of the other. Even though the problems seems to be separate, one cannot solve  the first problem without having to solve the second. In other words, after the creation of the two-way communication in the \glsreset{wsn}\gls{wsn} which adopts \gls{ipv6} over \gls{6lowpan}, other networks can only support \gls{ipv4} and therefore the communication would not be possible, and this is due to the heterogeneous nature of the system. This is why the two-way communication was created, but only after solving the heterogeneity problem, everything worked.
The system is divided into four subsystems:
\begin{itemize}
\item The fully \gls{ipv6} supported \gls{wsn}
\item The heterogeneity handler and middleman gateway server
\item The middle-ware
\item The mobile client
\end{itemize}
\paragraph{}
Each of these four components is there to solve part of the problem. Each of these subsystems will be discussed extensively in Chapter 4 from the architectural aspect of the component, and in Chapter 5 from a deployment point of view. 
\paragraph{}
It is important to note that some of the components will be discussed from the architectural point in Chapter 4 but not from an implementation aspect in Chapter 5. The sole reason behind this is the fact that some of the equipment required for this thesis were not on time and therefore the parts in the implementation requiring such equipment were omitted in order to meet the deadlines.
\section{Outcome of the Study}
\paragraph{}
The outcome of the thesis is to design and implement a test-bed that is able to show the two-way communication in action. It will also have to show its integration into the \glsreset{iot}\gls{iot} by showing that every node in the system can communicate autonomously and with any other node in the Internet. Another \gls{iot} concept is the ability of any node in the \gls{wsn} to consume and produce information that is communicated through the network.
\paragraph{}
The autonomous production of information is shown by the fact that every sensor node in the \gls{wsn} reads sensory data from its sensors, that are attached to appliance, in order to sense their consumption. This information about consumption is produced and communicated periodically. The consumption of data is proved by showing the execution of the commands communicated through the network by external nodes in the \gls{iot}. These commands request the sensor node to turn On or Off an appliance.
\paragraph{}
The test-bed should also show that communication between any two nodes in the \gls{iot} is successful and reliable, and that the heterogeneity of the system is transparent to the nodes in the system. This was achieved by creating a software tool deployed in the gateway which is between the \gls{wsn} and the outside world. This software tool is perceived as a heterogeneity handler, and also a communication middleman that translates the communication to the appropriate technology needed by the recipient.
\paragraph{}
The testbed will be able to show the two-way communication in action and will also show that the heterogeneity is not a barrier.

@@

Once the data is received from a client, it assumes that there are two data items separated by a space. The first item refers to the led number(1,2,3) and the second refers to the command 0 meaning turn Off and 1 means turn On. The string is tokenized and the data items are extracted and the command is executed accordingly.

@@

\chapter{Findings}
\paragraph{}
The system was built to monitor and control home appliances, and to also provide an infrastructure for future platforms to be integrated in this system. An example of infrastructures that can be the integrated is the \glsreset{dr}\gls{dr}. To build overlying system's, one should make sure that the system achieves a performance level that one can count on. The purpose of this chapter is to measure the performance of the system. It is done by conducting two experiments which will help us gather data, and based on that, an interpretation of this data will be done to verify if the system shows an acceptable level of performance.
\section{Experimentations}
\paragraph{}
To measure the performance of the system, two experiment were conducted where the goal is to quantify the performance of the system. There is an architectural decision madein \gls{6lowpan}, and we would like to see if it add more performance to the system. The architectural decision is the following: There is more priority given to routed packet that the processed ones. As a result, this clearly pushes us to think that the delay of the system will not follow the same growth trend as the network traffic.
\paragraph{Experiment 1:}
We start with one mote in the \gls{wsn} that will be interacting with the outside world. We then start sending ping command periodically (once every 1 second) and record the delay and the jitter. Once every two minutes, we will add another mote to the \gls{wsn} and continue pinging the same mote. We keep adding motes every 2 minutes until we reach 6 motes which is the maximum number of motes available. This will enable us to collect data about the mote delay and jitter as we add more traffic to the system.
\paragraph{Experiment 2: }
The second experience is meant to measure the time it takes to transform the packet. In other words, measure the extra delay and jitter when having the Gateway Transformation Process as part of the system. The timer starts before the sniffing of the packet and ends right after the packet is transformed and communicated.
We believe that this transformation adds delay, but we are hoping that this extra delay is not significant enough to affect the whole system.
\section{Data Gathering and Analysis}
\paragraph{}
After both experiments were conducted, we collected a significant amount of data that was used to measure the performance. In addition, we analyze and interpret the results to make judgments about the overall performance of the system.
\subsection{Experiment 1}
\paragraph{}
In the first experiment, the same mote was pinged once every second and we added new motes to the system every two minutes. The experiment as a whole took 12 minutes or 720 pings. Table \ref{table:exp1} shows the result of the experiment.

@@

This project enabled us to create an infrastructure that would be part of the \glsreset{iot}\gls{iot} and move towards having the smart home system's integrated within the Smart Grid. At this stage and counting solely on this work, one cannot clearly see the integration of smart home systems within the \gls{iot}, but this is an infrastructure that provides two main services which are the establishment of the two-way communication between the smart home devices and the outside world, and the heterogeneity handler. The two-way communication enables the \glsreset{wsn}\gls{wsn} to monitor home appliances and send sensor data to a repository and this last can be used to simulate a real world \gls{ami}. The second direction in this communication gives the ability for clients to control appliances by turning them On or Off.

@@

\subsection{Install Software in Motes}
The first thing to do is to install the backstation within the Mote sink (the mote with the \gls{usb} IO) To install it we do the following:
\begin{enumerate}
\item Plug the mote and type the command dmesg to see what \gls{usb} ID was assigned to it. It will generally be USB0.
\item cd \$TOSROOT/apps/IPBaseStation
\item make micaz blip install  mib520, /dev/ttyUSB0
\end{enumerate}
The backstation is now installed. We will now install the Mote TCP Client and Server in each sensor node. To do it, we should first download the code from the SVN repository.
\begin{enumerate}
\item Make sure you have SVN by typing svn --version
\item If you do not have SVN type sudo apt-get install subversion
\item download the file:
\item cd
\item svn checkout http://grid-aui-tinyos.googlecode.com/svn/trunk/ grid-aui-tinyos
\item cd grid-aui-tinyos//programming/TCPEcho
\end{enumerate}
Now take the MIB520 programming card, plug it and type dmesg to see where it plugged. lets call the location *USB. Now plug each mote in the MIB520 card and for each plugged mote do the following:

>>>>

\usepackage{etoolbox}
\makeatletter
\patchcmd{\thebibliography}{%
  \chapter*{\bibname}\@mkboth{\MakeUppercase\bibname}{\MakeUppercase\bibname}}{%
  \chapter{References}}{}{}
\makeatother

@@

\chapter{Abstract}
\paragraph{}
Information Technology is growing exponentially and largely impacting our daily lives. The current evolution in computing is the \gls{iot}. It is smoothly evolving from an Internet of People to and Internet of \textit{Things}. By 2020, the Internet is expected to have 50 billion \textit{Things} connected to it. 
\paragraph{}
However, such a migration induces a strong level of complexity when handling interoperability between heterogeneous \textit{Things}, e.g. Wireless Sensors, RFID, and mobile devices.
\paragraph{}
Another migration is taking place in the power grid that is evolving towards the \gls{sg}. In fact, Information Technology will play a key role in implementing needed functions such as \gls{dr}\glsreset{dr} and integrating Distributed and Renewable Energy Sources into \gls{sg}. The \gls{iot} is expected to enable a smooth integration of Distributed Renewable Energies within the electric grid.
\paragraph{}
This thesis focuses on the development of an \gls{iot} test-bed, using wireless sensors, that will be used for Smart Grid applications. The test-bed built provides an IPv6 gateway, remote control of wireless sensor nodes which is built on top of \gls{6lowpan}\glsreset{6lowpan}, and mobile access to the system.
\paragraph{}
This study results in the creation of a test-bed that provides an infrastructure for integrating \gls{sg} into \gls{iot} and building overlying services such as \gls{dr}. Therefore, the system's performance was put to test by conducting two experiments where the delay and jitter were measured to make sure the system can handle overlying services in future projects.

\chapter*{Résumé}
\paragraph{}
Les nouvelles technologies de l'information connaissent une croissance exponentielle et prennent de plus en plus d'impact sur nos vies. L'évolution en cours de l'informatique est \glsreset{iot}\gls{iot}. L'Internet est discrètement en train de transitioner d'un réseau pour les personnes vers un réseau pour les \textit{Choses}. En 2020, il est prévu d'avoir 50 milliards \textit{Choses} connectées à Internet.
\paragraph{}
Cependant, une telle migration induit un niveau élevé de complexité lors de la manipulation de l'interopérabilité entre \textit{Choses} hétérogènes, par exemple, Capteurs sans fil, RFID, et appareils mobiles.
\paragraph{}
Une autre migration commence dans le réseau électrique qui est entrain de se transformer en un nouveau réseau appelé \glsreset{sg}\gls{sg}. Ce nouveau réseau compte principalement sur les avantages des technologies de l'information qui lui permettra d'intégrer de nouvelles fonctionnalités telle que \glsreset{dr}\gls{dr} et la l'exploitation décentralisée des énergies renouvelables. Pour cela, \gls{iot} va permettre une douce intégration des énergies renouvelables dans le réseau électrique.
\paragraph{} 
Cette thèse porte à développer un \gls{iot} prototype qui utilise des capteurs sans fil pouvant être utilisés dans le Smart Grid. Ce prototype fournit des services tels qu'une passerelle IPv6, le contrôle à distance des capteurs sans fil, et un accès mobile au système.
\paragraph{}
Cette étude conduit à la création d'un prototype qui propose une infrastructure for intégrer \gls{sg} au sein de \gls{iot} et donc permettre de développer des services tel que \gls{dr}. Pour cela, les performances du système doivent être mises en test pour s'assurer de la possibilités de compter sur ce prototype. Dans ce contexte, deux expériences ont été menées mesurant le retard et le gigue pour savoir si le système peut supporter d'autres applications.
\setcode{utf8}

@@

\section{Importance of Study}
\paragraph{}
The global electrical grid is the basis that gives us this quality of life that most people around the world enjoy. The global electrical grid is sometimes called the largest machine ever made by man \cite{ref1}. It is a widespread and distributed machine that provides electricity to households, companies and organizations. Therefore the whole world's economy depends on it. The electrical grid knows numerous problems such as blackouts, and current instabilities that cost billions of dollars yearly to both the utilities and insurances.
\paragraph{}
The global energy demand is significantly growing and the sources of energy such as fossil energies are becoming scarce. Therefore, there is a need to move from exploiting fossil energy sources to using renewable energies such as Solar energy, wind energy and thermal energy. In addition, these renewable energies should be distributed throughout the power grid and permit to have less energy transportation losses.
\paragraph{}
The \glsreset{sg}\gls{sg} comes as a solution to the weaknesses of the traditional electrical grid. \gls{sg} is an intelligent, auto-balancing, self-monitoring, and self-healing electrical grid \cite{ref2} whose energy sources are distributed between utilities and consumers. The nature of the \gls{sg} as having distributed energy sources, reduces numerous losses that arise from the long distance transmission lines, and also from the stress and the overload of the electrical lines.
\paragraph{}
Another revolution began, and it is taking place in Information Technology. It is called the \glsreset{iot}\gls{iot}. \gls{iot} is the era of autonomous machine being able to produce and consume information, and being able to act autonomously without the guidance of humans, to perform their tasks. Cisco predicts that the number of devices to be connected to the Internet in 2020 will break the 50 billions devices from 500 million in 2003. That would make 6.58 devices connected for each person in the world whereas that same ratio was 0.07 17 years earlier \cite{ref3}. This clearly shows that the shift to the \gls{iot} will be exponential.
\paragraph{}
\glsreset{sg}\gls{sg} should benefit from the \gls{iot}. This integration constitutes the main topic of this study and goes along with increasing interest recently raised in both academia and industry, on the inherent correspondence between \gls{sg} and \gls{iot}. This  correspondence stems mainly from the need to disseminate information between the \gls{sg} components in order to make optimal decisions about energy use. This information dissemination could only be optimal when deploying ubiquitous sensors and meters, and using the available network infrastructure, i.e. the Internet.

\section{Rationale of Study}
\paragraph{}
This project has been done in order to investigate ways the \gls{iot} can benefit the \gls{sg}. The Smart Grid needs a set of novel functionalities. One of these is the two-way communication that will make it possible for the utility to communicate with consumers and vice versa. This is possible in the Internet Of Things as devices in \gls{iot} communicate by producing and consuming information. Numerous ideas assume the existence of the two-way communication such as the \gls{dr} which offers a set of services for both the power provider and the consumer.
\paragraph{}
The \glsreset{iot}\gls{iot} uses different technologies such as \gls{wsn}, \gls{rfid}, \gls{ipv6} and \gls{6lowpan}\glsreset{6lowpan} which makes the Internet of Things a very heterogeneous and complex system \cite{ref4}. One of the challenges is to deal with this heterogeneity in order to make the communication smooth and transparent to the parties involved. This heterogeneity should become invisible to any two devices willing to communicate in the \gls{iot}.

\section{Problem Statement}
\paragraph{}
The \glsreset{sg} \gls{sg} relies on the two-way communication to build a whole communication framework and a protocol stack namely the \glsreset{dr}\gls{dr} and other services built on top of this last. From the consumer's/household's perspective or more precisely the household's side,there are smart energy management systems in the market nowadays, but none of them offers an integration with \gls{sg}. Such an integration through the two-way communication will to make it possible for any appliance to report its real-time energy consumption, and to be controlled remotely. Such pillar is missing for the time being, and therefore the integration of the smart home within the smart grid is not doable. Also, in order for \gls{sg} to be integrated into the \gls{iot}, all devices should be able to communicate in the two ways.
\paragraph{}
\gls{sg} counts on various technologies to achieve its main goals. The use of different technologies comes at a cost which is the heterogeneous nature of the system. As numerous components count on different technologies, there is a need to interface them in order to make this heterogeneity transparent to the all components of the system. As a matter of fact, a software should be built within the system whose main goal is to take care of making the communication technology-transparent. This component will give the possibility to all components of the system to work hand-in-hand in order to achieve their goals independently of the data link layer (Ethernet, Wifi or Zigbee) or the network layer technology (\gls{ipv4} or \gls{ipv6}) they are using.

\section{Purpose of Study}
 \paragraph{}
The goal of the study is create the two-way communication, and solve the heterogeneity that arises from integrating smart homes with the smart grid. In fact, each components in the smart home will become part of the Internet of Things by making it communicate with any device in the Internet by having an \gls{ipv6} address using \gls{6lowpan} whose goal is to provide the TCP/IP to the smallest devices and also those with the least computational resources.
\paragraph{}
The first purpose of this study is to create a testbed with the two-way communication. To do so, \gls{6lowpan} would had to be integrated within every device in the \gls{wsn}. This would make every sensor node (mote) identifiable in the network, ready to communicate and be part of the \gls{iot}. This communication will go in both directions and based on \gls{6lowpan}; the two-way communication will be built.
\paragraph{}
The second purpose of this study is to hide the heterogeneity that arises from the integration of smart homes into the grid and also when making the whole system part of the \glsreset{iot}\gls{iot}. To solve such a problem, we add an extra component that will track all  packets and change them to fit the standards of the targeted component.

\section{Scope of Study}
\paragraph{}
A variety of tools and technologies were used for this study.The two problems could not be solved one independently of the other. Even though the problems seems to be separate, one cannot solve  the first problem without having to solve the second. In other words, after the creation of the two-way communication in the \glsreset{wsn}\gls{wsn} which adopts \gls{ipv6} over \gls{6lowpan}, other networks can only support \gls{ipv4} and therefore the communication would not be possible, and this is due to the heterogeneous nature of the system. This is why the two-way communication was created, but only after solving the heterogeneity problem, everything worked.
The system is divided into four subsystems:
\begin{itemize}
\item The fully \gls{ipv6} supported \gls{wsn}
\item The heterogeneity handler and middleman gateway server
\item The middle-ware
\item The mobile client
\end{itemize}
\paragraph{}
Each of these four components is there to solve part of the problem. Each of these subsystems will be discussed extensively in Chapter 4 from the architectural aspect of the component, and in Chapter 5 from a deployment point of view. 
\paragraph{}
It is important to note that some of the components will be discussed from the architectural point in Chapter 4 but not from an implementation aspect in Chapter 5. The sole reason behind this is the fact that some of the equipment required for this thesis were not on time and therefore the parts in the implementation requiring such equipment were omitted in order to meet the deadlines.
\section{Outcome of the Study}
\paragraph{}
The outcome of the thesis is to design and implement a test-bed that will be integrated into the \glsreset{iot}\gls{iot} by showing that every node in the system can communicate autonomously and with any other node in the Internet. Another \gls{iot} concept is the ability of any node in the \gls{wsn} to consume and produce information that is communicated through the network.
\paragraph{}
The autonomous production of information is shown by the fact that every sensor node in the \gls{wsn} reads sensory data from its sensors, that are attached to appliance, in order to sense their consumption. This information about consumption is produced and communicated periodically. The consumption of data is proved by showing the execution of the commands communicated through the network by external nodes in the \gls{iot}. These commands request the sensor node to turn On or Off an appliance.
\paragraph{}
The test-bed should also show that communication between any two nodes in the \gls{iot} is successful and reliable, and that the heterogeneity of the system is transparent to the nodes in the system. This was achieved by creating a software tool deployed in the gateway which is between the \gls{wsn} and the outside world. This software tool is perceived as a heterogeneity handler, and also a communication middleman that translates the communication to the appropriate technology needed by the recipient.
\paragraph{}
The testbed will be able to show the two-way communication in action and will also show that the heterogeneity is not a barrier.

@@

Once the data is received from a client, it assumes that there are two data items separated by a space. The first item refers to the led number(1,2,3) and the second refers to the command 0 meaning turn Off and 1 means turn On. The string is tokenised and the data items are extracted and the command is executed accordingly.

@@

\chapter{Findings}
\paragraph{}
The system was built to monitor and control home appliances, and to also provide an infrastructure for future platforms to be integrated in this system. An example of infrastructures that can be the integrated is the \glsreset{dr}\gls{dr}. To build overlying system's, one should make sure that the system achieves a performance level that one can count on. The purpose of this chapter is to measure the performance of the system. It is done by conducting two experiments which will help us gather data, and based on that, an interpretation of this data will be done to verify if the system shows an acceptable level of performance.
\section{Experimentations}
\paragraph{}
To measure the performance of the system, two experiment were conducted where the goal is to quantify the performance of the system. There is an architectural decision made in \gls{6lowpan}, and we would like to see if it add more performance to the system. The architectural decision is the following: There is more priority given to routed packet that the processed ones. As a result, this clearly pushes us to think that the delay of the system will not follow the same growth trend as the network traffic.
\paragraph{Experiment 1:}
We start with one mote in the \gls{wsn} that will be interacting with the outside world. We then start sending ping command periodically (once every 1 second) and record the delay and the jitter. Once every two minutes, we will add another mote to the \gls{wsn} and continue pinging the same mote. We keep adding motes every 2 minutes until we reach 6 motes which is the maximum number of motes available. This will enable us to collect data about the mote delay and jitter as we add more traffic to the system.
\paragraph{Experiment 2: }
The second experience is meant to measure the time it takes to transform the packet. In other words, measure the extra delay and jitter when having the Gateway Transformation Process as part of the system. The timer starts before the sniffing of the packet and ends right after the packet is transformed and communicated.
We believe that this transformation adds delay, but we are hoping that this extra delay is not significant enough to affect the whole system.
\section{Data Gathering and Analysis}
\paragraph{}
After both experiments were conducted, we collected a significant amount of data that was used to measure the performance. In addition, we analyze and interpret the results to make judgments about the overall performance of the system.
\subsection{Experiment 1}
\paragraph{}
In the first experiment, the same mote was pinged once every second and we added new motes to the system every two minutes. The experiment as a whole took 12 minutes or 720 pings. Table \ref{table:exp1} shows the result of the experiment.

@@

This project enabled us to create an infrastructure that would integrated into \glsreset{iot}\gls{iot} and move towards having the Smart Home systems' integrated within the Smart Grid. At this stage and counting solely on this work, one cannot clearly see the integration of smart home systems within the \gls{iot}, but this is an infrastructure that provides two main services which are the establishment of the two-way communication between the smart home devices and the outside world, and the heterogeneity handler. The two-way communication enables the \glsreset{wsn}\gls{wsn} to monitor home appliances and send sensor data to a repository and this last can be used to simulate a real world \gls{ami}. The second direction in this communication gives the ability for clients to control appliances by turning them On or Off.

@@

\subsection{Install Software in Motes}
The first thing to do is to install the back station within the Mote sink (the mote with the \gls{usb} IO) To install it we do the following:
\begin{enumerate}
\item Plug the mote and type the command dmesg to see what \gls{usb} ID was assigned to it. It will generally be USB0.
\item cd \$TOSROOT/apps/IPBaseStation
\item make micaz blip install  mib520, /dev/ttyUSB0
\end{enumerate}
The backstation is now installed. We will now install the Mote TCP Client and Server in each sensor node. To do it, we should first download the code from the SVN repository.
\begin{enumerate}
\item Make sure you have SVN by typing svn --version
\item If you do not have SVN type sudo apt-get install subversion
\item download the file:
\item cd
\item svn checkout http://grid-aui-tinyos.googlecode.com/svn/trunk/ grid-aui-tinyos
\item cd grid-aui-tinyos//programming/TCPEcho
\end{enumerate}
Now take the MIB520 programming card, plug it and type dmesg to see where it plugged. lets call the location *USB. Now plug each mote in the MIB520 card and for each plugged mote do the following:
